Clustering time series based on probability distributions across temporal granularities

### 2021-11-17 main.Rmd
adding smart meter data and Hennig paper reference

Determining the number of clusters is typically a difficult task. Many metrics have been defined for choosing clusters. Most metrics for choosing the optimal number of clusters, are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters. Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy] and the separation index ($sindex$) [@Hennig2014-ah]. These are constructed by balancing within-cluster homogeneity and between-cluster separation.

@@

Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO [@make-these-two-URLs-into-proper-citations]. The full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual. We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation. Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying starting and end dates.  Additionally, there were missing values in many series. (The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure is related to granularities.

>>>>

Determining the number of clusters is typically a difficult task. Many metrics have been defined for choosing clusters. Most metrics for choosing the optimal number of clusters, are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters. Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy] and the separation index ($sindex$) [@Hennig2019-eq; @Hennig2014-ah]. These are constructed by balancing within-cluster homogeneity and between-cluster separation.

@@

Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from @smart-meter. The full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual. We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation. Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying starting and end dates.  Additionally, there were missing values in many series. (The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure is related to granularities.
